---
title: "Tuna Demand Analysis"
author: "Group 5"
date: "`r Sys.Date()`"
output: html_document
---

```{r Packages, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(tidyverse)
library(lubridate)
library(tsibble)
```


```{r Read Data, message=FALSE}
demand <- read_csv("Tuna_demand.csv")
price <- read_csv("Tuna_price.csv")
cpi <- read_csv("PCE_CPI.csv")
```


```{r Data Prep}

# date cleansing
demand <- demand %>%
  mutate(date = as.Date(as.character(demand$WeekEndingDay), format = "%Y%m%d")) %>%
  select(-c(WeekEndingDay)) %>%
  relocate(date) %>%
  pivot_longer(!date, names_to = "brand", values_to = "demand")

price <- price %>%
  mutate(date = as.Date(as.character(price$WeekEndingDay), format = "%Y%m%d")) %>%
  select(-c(WeekEndingDay)) %>%
  relocate(date) %>%
  pivot_longer(!date, names_to = "brand", values_to = "price")

```


```{r Prep 2}
# join
tuna <- demand %>%
  inner_join(price)

# covid
tuna <- tuna %>% 
  mutate(Covid = ifelse(date >= '2020-01-01', 1, 0))
tuna$Covid <- as.factor(tuna$Covid)
```




# ANALYSIS

1. Regress demand on price and brand. Since brand variable is categorical, it should be a factor

```{r}
summary(lm(demand ~ price + as.factor(brand), tuna))
```

Regression equation: demand = 4,010,270 - 1,513,394 * price + 296,010 * I(brand=Others) + 664,070 * I(brand=StarKist)

* I(Brand = xx) equals 1, if Brand is xx, and 0, otherwise.
  * when the brand is StarKist, I(brand=StarKist) = 1 and I(brand=Others) = 0
  
* Interpretation of the equation is that …
  * demand decreases by 1,513,394 when price increases by $1,
  * Other brands, on average, sold 296,010 units per week MORE than BumbleBee, and
  * StarKist, on average, sold 664,070 units per week MORE than Brand BumbleBee

* The p-value of the coefficients (in the 4th column in Coefficients table) implies that the price effect is significant.



1.a Regression with covid 

```{r}
summary(lm(demand ~ price + as.factor(brand) + as.factor(Covid), tuna))
```

Covid seems to be significant - demand goes up 583,346 during covid. People can't eat out, grocery shopping and dining in more? Tuna for lunch while working from home?


2. Month Sales Analysis

```{r Monthly data prep}
monthlyTuna = tuna %>%
  mutate(year = year(date), month = month(date, label = TRUE, abbr = FALSE)) %>%
  group_by(year, month, brand) %>%
  summarise(demand = sum(demand), price = mean(price))

monthlyTuna
```

```{r}
cpi <- cpi %>%
  mutate(date = as.Date(with(cpi, paste(Year, Month, '01',sep="-")))) %>%
  mutate(year = year(date), month = month(date, label = TRUE, abbr = FALSE)) %>%
  group_by(year, month) %>%
  select(-c(Year, Month, date)) %>%
  relocate(month, year)
```

```{r Monthly data prep}
# join
tuna_cpi_monthly <- monthlyTuna %>%
  inner_join(cpi)

# covid
tuna_cpi_monthly <- tuna_cpi_monthly %>% 
  mutate(Covid = ifelse(year >= 2020, 1, 0))
tuna_cpi_monthly$Covid <- as.factor(tuna_cpi_monthly$Covid)

```

Monthly regression

```{r}
summary(lm(demand ~ price + as.factor(brand), tuna_cpi_monthly))

summary(lm(demand ~ price + as.factor(brand) + as.factor(Covid), tuna_cpi_monthly))

summary(lm(demand ~ price + as.factor(brand) + CPI_U_food, tuna_cpi_monthly))

summary(lm(demand ~ price + as.factor(brand) + as.factor(Covid) + CPI_U_food, tuna_cpi_monthly))
```

Best adjusted R^squared 

lm(formula = demand ~ price + as.factor(brand) + as.factor(Covid), 
    data = tuna_cpi_monthly)



# Seasonality

```{r}
tuna_seas <- tuna %>% 
  mutate(week_of_year = week(date), month_of_year = month(date))
```




Basic Model w/ brand

```{r}
summary(lm(log(demand) ~ log(price) + as.factor(brand), tuna_seas))
```


```{r}
summary(lm(log(demand) ~ log(price) + as.factor(brand) + as.factor(month_of_year), tuna_seas))
```


Cold months seem to be significant - tuna demand goes down


```{r}
summary(lm(log(demand) ~ log(price) + as.factor(brand) + as.factor(month_of_year) + as.factor(Covid) + CPI_, tuna_seas))
```




# Log model

```{r}
summary(lm(log(demand) ~ log(price) + as.factor(brand), tuna))
```

The estimate of the log price coefficient is -1.346, i.e., the tuna demand is price elastic.








