---
title: "Exercise 1"
author: "Carter Rudek"
date: "`r Sys.Date()`"
output: html_document
---

# Lab material 1 – Log-log demand models and price elasticities

* This exercise is based on K. Wertenbroch (2006) “Measuring Price Promotion Effects: An Econometric Exercise in Measuring the Impact of Marketing Decision Making.”
* If you’ve not yet installed R, RStudio, and tidyverse package, install them by following directions in Application installation guide for BUSML7247.pdf file.
* In this material, what you have to type in R is highlighted in yellow.


```{r}
library(tidyverse)
```

```{r}
Data = read_csv("PricePromotionEffectsData.csv")
```

3. Basic model
  3.1. For each brand (X and Y), generate a contingency table for groups by class (lite and regular) and deal_X/deal_Y (whether the UPC was price-promoted or not). Summarize the number of observations (using n function in summarise function; see Chapter 3 of “R for Data Science” book (Chapter 5 in the web version) for the details of the functions), the average sales amount (oz_X/oz_Y), and the average price (pX/pY).


```{r}
Data %>% group_by(class, deal_X) %>%
  summarise(n = n(), avg_demand = mean(oz_X), avg_price = mean(pX))
```

```{r}
Data %>% group_by(class, deal_Y) %>%
  summarise(n = n(), avg_demand = mean(oz_Y), avg_price = mean(pY))
```


* For brand X,
  * The average price w/ the price promotion (0.0225 for lite, 0.0228 for regular) is lower than the average price w/o the price promotion (0.0298 for lite, 0.0297 for regular). This is obvious and there seems no difference between the classes (0.0225 vs. 0.0228 and 0.0298 vs. 0.0297).
  * The average sales w/ the price promotion (25455 for lite, 24125 for regular) is much greater than (about 9-10 times as great as) the average sales w/o the price promotion (2686 for lite, 2588 for regular). There seems no significant difference between the classes (25455 vs. 24125 and 2686 vs. 2588).
  * In sum, the table shows clear evidence of the price promotion effect for brand X.
* Brand Y shows evidence similar to Brand X.



3.2. Regress the log sales amount (log(oz_X)/log(oz_Y)) on the log price (log(pX)/log(pY)) as follows (the regression result is stored and will be used to
draw the demand curves):

```{r}
reg_X = lm(log(oz_X) ~ log(pX), Data)
reg_Y = lm(log(oz_Y) ~ log(pY), Data)
summary(reg_X)
```

```{r}
summary(reg_Y)
```

* The above regression results confirm the evidence above: the coefficient of the log price is -6.81 (brand X) and -6.48 (brand Y), implying that both brands are price elastic.

3.3. Generate demand curves as follows:

```{r}
alphaX = reg_X$coefficients[1]
betaX = reg_X$coefficients[2]
priceX = seq(min(Data$pX), max(Data$pX), 0.001)
demandX = exp(alphaX + betaX * log(priceX))
alphaY = reg_Y$coefficients[1]
betaY = reg_Y$coefficients[2]
priceY = seq(min(Data$pY), max(Data$pY), 0.001)
demandY = exp(alphaY + betaY * log(priceY))
plot(priceX, demandX, type = "l", col = "blue", xlab = "price", ylab = "demand", main = "Demand curve: basic model")
points(priceY, demandY, type = "l", col = "red")
legend("topright", legend = c("X", "Y"), col = c("blue", "red"), lty = 1)
```


* For brand X,
  * alphaX and betaX store the regression coefficient estimates of intercept and log price, respectively.
  * priceX stores a series of numbers from the observed minimum price of the brand (min(Data$pX)) to the observed maximum price of the brand (max(Data$pY)) at intervals of .001. seq function is used to generate a series like this. Refer to R help (type ? seq to see the help) for the details of the function.
 * Demand prediction at each price in priceX is calculated (exp(alphaX + betaX*log(priceX))) and stored in demandX variable.

* For brand Y, priceY and demandY are generated in the same manner.
* plot, points, and legend functions draw the demand curve above. Refer to R help for the details of the functions.
* Because the regression coefficient estimates of both brands are similar to each other (-16.05 vs. -14.82 for intercept; -6.81 vs. -6.48 for log price), the demand curves of both brands are similar to each other.









