---
title: "Exercise 1"
author: "Carter Rudek"
date: "`r Sys.Date()`"
output: html_document
---

# Lab material 1 – Log-log demand models and price elasticities

* This exercise is based on K. Wertenbroch (2006) “Measuring Price Promotion Effects: An Econometric Exercise in Measuring the Impact of Marketing Decision Making.”
* If you’ve not yet installed R, RStudio, and tidyverse package, install them by following directions in Application installation guide for BUSML7247.pdf file.
* In this material, what you have to type in R is highlighted in yellow.


```{r}
library(tidyverse)
```

```{r}
Data = read_csv("PricePromotionEffectsData.csv")
```

3. Basic model
  3.1. For each brand (X and Y), generate a contingency table for groups by class (lite and regular) and deal_X/deal_Y (whether the UPC was price-promoted or not). Summarize the number of observations (using n function in summarise function; see Chapter 3 of “R for Data Science” book (Chapter 5 in the web version) for the details of the functions), the average sales amount (oz_X/oz_Y), and the average price (pX/pY).


```{r}
Data %>% group_by(class, deal_X) %>%
  summarise(n = n(), avg_demand = mean(oz_X), avg_price = mean(pX))
```

```{r}
Data %>% group_by(class, deal_Y) %>%
  summarise(n = n(), avg_demand = mean(oz_Y), avg_price = mean(pY))
```


* For brand X,
  * The average price w/ the price promotion (0.0225 for lite, 0.0228 for regular) is lower than the average price w/o the price promotion (0.0298 for lite, 0.0297 for regular). This is obvious and there seems no difference between the classes (0.0225 vs. 0.0228 and 0.0298 vs. 0.0297).
  * The average sales w/ the price promotion (25455 for lite, 24125 for regular) is much greater than (about 9-10 times as great as) the average sales w/o the price promotion (2686 for lite, 2588 for regular). There seems no significant difference between the classes (25455 vs. 24125 and 2686 vs. 2588).
  * In sum, the table shows clear evidence of the price promotion effect for brand X.
* Brand Y shows evidence similar to Brand X.



3.2. Regress the log sales amount (log(oz_X)/log(oz_Y)) on the log price (log(pX)/log(pY)) as follows (the regression result is stored and will be used to
draw the demand curves):

```{r}
reg_X = lm(log(oz_X) ~ log(pX), Data)
reg_Y = lm(log(oz_Y) ~ log(pY), Data)
summary(reg_X)
```

```{r}
summary(reg_Y)
```

* The above regression results confirm the evidence above: the coefficient of the log price is -6.81 (brand X) and -6.48 (brand Y), implying that both brands are price elastic.

3.3. Generate demand curves as follows:

```{r}
alphaX = reg_X$coefficients[1]
betaX = reg_X$coefficients[2]
priceX = seq(min(Data$pX), max(Data$pX), 0.001)
demandX = exp(alphaX + betaX * log(priceX))
alphaY = reg_Y$coefficients[1]
betaY = reg_Y$coefficients[2]
priceY = seq(min(Data$pY), max(Data$pY), 0.001)
demandY = exp(alphaY + betaY * log(priceY))
plot(priceX, demandX, type = "l", col = "blue", xlab = "price", ylab = "demand", main = "Demand curve: basic model")
points(priceY, demandY, type = "l", col = "red")
legend("topright", legend = c("X", "Y"), col = c("blue", "red"), lty = 1)
```


* For brand X,
  * alphaX and betaX store the regression coefficient estimates of intercept and log price, respectively.
  * priceX stores a series of numbers from the observed minimum price of the brand (min(Data$pX)) to the observed maximum price of the brand (max(Data$pY)) at intervals of .001. seq function is used to generate a series like this. Refer to R help (type ? seq to see the help) for the details of the function.
 * Demand prediction at each price in priceX is calculated (exp(alphaX + betaX*log(priceX))) and stored in demandX variable.

* For brand Y, priceY and demandY are generated in the same manner.
* plot, points, and legend functions draw the demand curve above. Refer to R help for the details of the functions.
* Because the regression coefficient estimates of both brands are similar to each other (-16.05 vs. -14.82 for intercept; -6.81 vs. -6.48 for log price), the demand curves of both brands are similar to each other.




# Exercise 1 Problems

Q1. In the data, another marketing variable is available: the feature (in-store) display (feat_X for brand X and feat_Y for brand Y).  This variable is a binary (0-1) variable that indicates whether UPC was advertised or put on in-store display during the week.  Usually, a feature display promotion comes with a price discount. Let's analyze the feature display effect.  Please follow the steps below and answer the questions:

  1.	Generate a contingency table of average demand and average price.  Because a feature display promotion comes with a price discount, we need to look at both the price promotion effect and the feature display effect at the same time.  So, group the data by the deal indicator (deal_X/deal_Y) and the feature indicator (feat_X/feat_Y) for each brand and generate the contingency table.

```{r}
# Y
Data %>% group_by(deal_Y, feat_Y) %>%
  summarise(n = n(), avg_demand = mean(oz_Y), avg_price = mean(pY))
```

```{r}
# X
Data %>% group_by(deal_X, feat_X) %>%
  summarise(n = n(), avg_demand = mean(oz_X), avg_price = mean(pX))
```


  2. Add the feature (in-store) display effect into the model in the lab material for each brand.  For example, you need to regress log(oz_X) on log(pX) and feat_X for brand X.


```{r}
#X
reg_X2 = lm(log(oz_X) ~ log(pX) + feat_X, Data)
summary(reg_X2)
```


```{r}
#Y
reg_Y2 = lm(log(oz_Y) ~ log(pY) + feat_Y, Data)
summary(reg_Y2)
```

3. Answer the following questions for each brand:
  * Are there any differences in the average demand and the average price between the observations with the feature display promotion and those without the promotion?  Provide empirical evidence for your answer.

Yes, for brand Y, the average demand increases from 2581 to 13343 with the feature display promotion and no price promotion. The average price decreases from .0303 to .0295 with the feature display and no price promotion. The average demand again, increases, with the feature display from 14118 to 37008 when on price promotion. The average price decreases from .0226 to .0217 with the feature display and the price promotion.

Brand X is the same as brand Y, the average demand increases from 2329 to 4106 with the feature display promotion and no price promotion. The average price decreases from .03 to .0283 with the feature display and no price promotion. The average demand again, increases, with the feature display from 15275 to 27581 when on price promotion. The average price decreases from .0244 to .022 with the feature display and the price promotion.

* Does the regression with the feature variable confirm your findings in the contingency table?  Provide empirical evidence for your answer.

Yes, the regression results do confirm what we found in the contingency tables. Price is elastic for both brand x and y. The coefficient for the display feature for brand x is positive, at .4160. The coefficient for the display feature is also positive, at 1.6324.


* What is the size of the effect (the feature display coefficient)?  Interpret the value of the coefficient. (hint: you can write down the demand equation with the coefficient estimates as: log(D) = α  +  β x log(P)  +  δ x feat, where α, β, and δ are the coefficient estimates of the intercept, the log price, and the feature, respectively.  This is equivalent to  D = eα  x  Pβ  x  eδ, if feat = 1; D = eα  x  Pβ, if feat = 0.  Then, what is the change that the feature variable makes?)

For brand x, the coefficient for the feature display is .4160. This can be interpreted as, log(sales) increases .42% when brand x has a feature display. 
For brand y, the coefficient for the feature display is 1.6324. This can be interpreted as, log(sales) increases 1.6% when brand y has a feature display. 

* Compare the log price coefficient of your regression to that of the basic model in the lab material.  Does inclusion of the feature display effect affect the price promotion effect?  Provide empirical evidence for your answer.  (Drawing a demand curve w/ and w/o the feature display would be helpful to answer this question.)

X – basic: -6.8103, new: -5.8874
Y – basic: -6.4786, new: -3.7390 

Yes, including the feature display variable in our regression increased the price coefficient in the new model compared to the basic model. However, both brands are still elastic.



Q2. Let's analyze the competition effect.  Please follow the steps below and answer the questions:

1.	Add competitor's log price and feature variables into the model in Q1 (the model for feature display effect).  For example, you need to regress log(oz_X) on log(pX), feat_X, log(pY), and feat_Y for brand X.

```{r}
#X
reg_X3 = lm(log(oz_X) ~ log(pX) + feat_X + log(pY) + feat_Y, Data)
summary(reg_X3)
```


```{r}
#Y
reg_Y3 = lm(log(oz_Y) ~ log(pY) + feat_Y + log(pX) + feat_X, Data)
summary(reg_Y3)
```

2.	Answer the following questions for each brand:

* Is there a competitor price effect on sales?  Is there a competitor feature display effect on sales?  Provide empirical evidence for your answer.

Yes, for both brands.

For Brand x, the log(pY) coefficient is positive at .001, meaning brand x sales go up as brand y price goes up. Oppositely, the coefficient for feat_y in the brand x model is negative at -.0297, meaning brand x sales go down when brand y has a feature display.

For Brand y, the log(px) coefficient is positive at .6424, meaning brand y sales go up as brand x price goes up. This is stronger than we saw for brand x. The coefficient for feat_x in the brand y model is also positive at .011, meaning brand y sales go up when brand x has a feature display. This is different than what we saw for brand x.

* What is the size of the competitor price effect (the competitor price coefficient)?  Interpret the value of the coefficient.  (hint: think about the cross-price elasticity.)

In the brand x model, the coefficient for log(pY) is 0.001089
In the brand y model, the coefficient for log(pX) is 0.64244

Whereas the coefficients for price of the brand is it’s own elasticity of demand, these coefficients are the cross-price elasticity of demand. Meaning, they are the % change in the brands sales if the price of the other brand increases by 1%

Brand x sales increase by .001% if brand y price increases by 1%. Brand y sales increase by .64% if brand x increases their price by 1%

* Compare the log price coefficient of your regression to that of the model for feature display effect in Q1.  Does inclusion of the competitor's variables affect the own-price promotion effect?  Provide empirical evidence for your answer.

In the original Q1 model for feature display, the coefficients were:
brand x log price: -5.8874
brand y log price: -3.7390

the coefficients when including competitor’s variables:
brand x log price: -5.896607
brand y log price: -3.76910

For both brands, the coefficients slightly decreased when adding competitor variables but this change does not seem significant.














