---
title: "Marketing W1 In Class"
author: "Carter Rudek"
date: "`r Sys.Date()`"
output: html_document
---



### Activity 1 lab material – Bass diffusion model

* This material is based on “Forecasting the Adoption of a New Product” by E. Ofek (2016).

* If you’ve not yet installed R, RStudio, and tidyverse package, install them by following directions in Application
installation guide for BUSML7247.pdf file.

* In this material, what you have to type in R is highlighted in yellow.

```{r}
library(tidyverse)
```

2. Data preparation
2.1. Read data (CellularSubscriberData.csv, available in Carmen) by using read_csv function as follows. This command reads the csv (comma separated values) file and store the information in a data frame named Data. Refer to Chapter 8 of “R for Data Science” book (Chapter 11 in the web version) for the details of read_csv function.

```{r}
Data = read_csv("CellularSubscriberData.csv")
```

* If you cannot read the data, check if the working directory is the same as the folder where you’ve saved the data file by using getwd function and correct the working directory by using setwd function.

2.2. Take a look at the data frame by typing the name of the data frame (Data) as follows. There are two variables – Year and CumSales (cumulative

```{r}
Data
```

* The data frame contains only the sales data from 1983 to 1996. We have extra four time periods (1997 through 2000) for prediction.

2.3. To fit the data to the Bass diffusion model, we additionally need sales amount at each period (denoted S(t) in “Forecasting the Adoption of a New Product” article), the lagged cumulative amount (denoted N(t-1) in “Forecasting the Adoption of a New Product” article), and the square

```{r}
Data = Data %>%
 mutate(CumSalesLag = lag(CumSales, default = 0)) %>%
  mutate(CumSalesLagSq = CumSalesLag ^ 2) %>%
  mutate(Sales = CumSales - CumSalesLag)
```

* The first mutate function generates N(t-1) by using lag function. Here, the default option in the lag function must be set to 0 because the lagged variable for the first time period would be NA without the default option. See Chapter 3 of “R for Data Science” book (Chapter 5 in the web version) and R help (type ? dplyr::lag to see the help) for the details of the function.

* The second mutate function generates N(t-1)2 by computing the lagged variable to the power of 2. ^ is the exponent operator.

* The third mutate function generates S(t) = N(t) – N(t-1).


2.4. Take a look at the data frame again. You can see all of the variables we need.

```{r}
Data
```

3. Fitting the data to the Bass diffusion model

3.1. The appendix of “Forecasting the Adoption of a New Product” article shows the simplest estimation method using OLS (ordinary least squares) regression. This method does not guarantee stable and robust estimates, but can be used without any advanced statistical techniques. There is an error in the equations in the appendix, and the following equations are correct ones.

* The Bass model can be expressed as the following:
S(t) = [p + (q/m) N(t-1)] [m – N(t-1)] = pm + (q-p) N(t-1) – (q/m) N(t-1)2
or, more simply:
S(t) = β0 + β1 N(t-1) + β2 N(t-1)2

* Implying that
β0 = pm
β1 = q – p
β2 = – q/m

* Solving the equations above with respect to p, q, and m, we have
m = [ – β1 – (β12 – 4β0β2)1/2 ] / (2β2) or [ – β1 + (β12 – 4β0β2)1/2 ] / (2β2)
p = β0/m
q = – mβ2

3.2. Run an OLS regression by using lm function as follows.

```{r}
ols_out = lm(Sales ~ CumSalesLag + CumSalesLagSq, Data)
```

* The first input of the function specifies the regression equation – S(t) = β0 + β1 N(t- 1) + β2 N(t-1)2 . For details, the left-hand side of the equation, i.e., dependent variable, is S(t), which is Sales variable. In the right-hand side of the equation, we have three independent variables – a constant, N(t-1), and N(t-1)2. The constant will be automatically added into the regression, so we need to specify only the other two variables, which are CumSalesLag and CumSalesLagSq. In the lm function, 4 the dependent variable is specified before the ~ symbol and the independent variables are listed after the ~ symbol with a separator (the + symbol).

* The second input simply specifies the name of the data frame where the data exist. The command above stores the OLS regression result in ols_out.


3.3. To see the result, take the summary of the result as follows. We will learn (or review) the summary of an OLS regression later.

```{r}
summary(ols_out)
```

3.4. We can access the coefficient values via ols_out$coefficients:

```{r}
beta0 = ols_out$coefficients[1]
beta1 = ols_out$coefficients[2]
beta2 = ols_out$coefficients[3]
```

3.5. Compute p, q, and m from the β’s (beta0, beta1, and beta2) as follows.

```{r}
m_1 = (-beta1 - sqrt(beta1 ^ 2 - 4 * beta0 * beta2)) / (2 * beta2)
m_2 = (-beta1 + sqrt(beta1 ^ 2 - 4 * beta0 * beta2)) / (2 * beta2)
m_1
```
```{r}
m_2
```

```{r}
m = m_1 # Replace m_2 for m_1 if m_2 is positive...
p = beta0 / m
q = -m * beta2
p
```

```{r}
q
```

* Note that there are two possible solutions of m (m_1 and m_2) and we need to take the positive one of them. In this exercise, m_1 is positive (92.50633) but m_2 is negative (-0.1250123), so m_1 is taken as the solution of m.

4. Generating prediction from the model parameters

4.1. The Bass model implies a dynamic relationship between the previous period data and the current period data. That is, prediction should be generated sequentially over time. To do so, create a variable (SalesPred) where the prediction will be stored and an auxiliary variable (N_lag) for prediction, as follows.

```{r}
Data = Data %>% mutate(SalesPred = 0)
N_lag = 0
```


4.2. Using a for loop, we can generate prediction (S_t) sequentially and store it in the variable (SalesPred) we’ve created:

```{r}
for(TIME in 1:nrow(Data)) {
  S_t = (p + q / m * N_lag) * (m - N_lag)
  Data$SalesPred[TIME] = S_t
  N_lag = N_lag + S_t
  }
```


4.3. Plot them! (See R help – type ? plot and ? points – for the details of the plotting functions)

```{r}
plot(Data$Year, Data$Sales, type = "p",
  ylim = c(0, 1.2 * max(Data$Sales, Data$SalesPred, na.rm = TRUE)),
  xlab = "Year", ylab = "Unit sales (in millions)")
  points(Data$Year, Data$SalesPred, type = "b",
  pch = 23, col = "red", lty = 2)
```

4.4. Generate the prediction of the cumulative sales (CumSalesPred) from the sales prediction (SalesPred) as follows.

```{r}
Data = Data %>% mutate(CumSalesPred = cumsum(SalesPred))
```

* cumsum function is useful to generate a cumulative aggregate. See Chapter 3 of “R for Data Science” book (Chapter 5 in the web version) for the details of the function.

4.5. Plot them!

```{r}
plot(Data$Year, Data$CumSales, type = "p",
  ylim = c(0, 1.2 * max(Data$CumSales, Data$CumSalesPred, na.rm =
TRUE)),
  xlab = "Year", ylab = "Cumulative sales (in millions)")
  points(Data$Year, Data$CumSalesPred, type = "b",
  pch = 23, col = "red", lty = 2)
```


### Question and Answers

###### Q1. In the lab, you may have p = 0.0007, q = 0.544, and m = 92.5 millions. Interpret those estimates of the model parameters:
* What does 0.0007 of p imply?

p in a Bass diffusion model is the innovation parameter, so 0.0007 of p means that, on average, 0.07% of non-adopters are estimated to adopt the service at each time period through external factors such as the effects of advertising and other mass media. This is quite small, implying that the service is estimated to have been adopted very slowly at first.

* What does 0.544 of q imply?

q in a Bass diffusion model is the imitation parameter, so 0.544 of q means that, on average, 54.4% of non-adopters are estimated to be influenced by others. This is high, implying that the service is estimated to rapidly grow once the adoptions take off.

* What does 92.5 millions of m imply?

This number indicates the estimated market potential. So, about 92.5 million people are estimated to finally adopt the service in the market.

###### Q2. Compared to the total number of mobile subscriptions in the US (404.58 millions, as of 2018), the estimate of m from our analysis (92.5 millions) seems to be too small to reflect reality. This is a typical example of Bass diffusion model's limitation. What do you think leads to the extremely under-estimated market potential?

This is a well-known limitation of Bass diffusion models: the models just extrapolate from the past trend of adoptions. The extrapolation can be valid only if the trend used for the extrapolation has been continuously found. But, the cell phone 20-30 years ago is a totally different animal from the current cell phones in terms of not only the service speed and quality but also what we can do with the cell phone. That is, the current trend is absolutely different from the trend used for the extrapolation, leading to the extremely under-estimated market potential.

This is an example of why a long-run forecast is very difficult. We must, at least partially, rely on extrapolations for a long-run forecast and it is necessary to assume that the trend we’ve observed continues… But, this assumption is very dangerous: think about what we’ve decided and what have happened with COVID-19.


###### Q3. Download the following csv data file (available in the assignment webpage)
* Data file: AssassinsCreed7Data.csv and fit the data to the Bass diffusion model. Here is useful information:

* The dataset contains weekly sales (not cumulative sales) of Assassin's Creed 7 video game for the first 15 weeks after release. Thus, you have to modify the R code of the data preparation part for estimation.

Report p, q, and m estimates with your own interpretation of the estimates (you don't need to draw a fancy table to do this. Just list the numbers...). Generate and visualize the prediction of the sales. Can you provide any managerial suggestion(s) from the predicted diffusion curve?

```{r}
# read data file
Data2 = read_csv("AssassinsCreed7Data.csv")
Data2 = Data2 %>%
  mutate(CumSales = cumsum(Sales)) %>%
  mutate(CumSalesLag = lag(CumSales, default = 0)) %>%
  mutate(CumSalesLagSq = CumSalesLag ^ 2)
```

```{r}
# run OLS
ols_out = lm(Sales ~ CumSalesLag + CumSalesLagSq, Data2)
summary(ols_out)
```

```{r}
# compute p, q, and m from the OLS coefficients
beta0 = ols_out$coefficients[1]
beta1 = ols_out$coefficients[2]
beta2 = ols_out$coefficients[3]
m_1 = (-beta1 - sqrt(beta1 ^ 2 - 4 * beta0 * beta2)) / (2 * beta2)
m_2 = (-beta1 + sqrt(beta1 ^ 2 - 4 * beta0 * beta2)) / (2 * beta2)
m_1
```
```{r}
m_2
```

```{r}
m = m_1 # Replace m_2 for m_1 if m_2 is positive...
p = beta0 / m
q = -m * beta2
p
```

```{r}
q
```


```{r}
Data2 = Data2 %>% mutate(SalesPred = 0)
N_lag = 0
for(TIME in 1:nrow(Data2)) {
  S_t = (p + q / m * N_lag) * (m - N_lag)
  Data2$SalesPred[TIME] = S_t
  N_lag = N_lag + S_t
}
plot(Data2$week, Data2$Sales, type = "p",
  ylim = c(0, 1.2 * max(Data2$Sales, Data2$SalesPred, na.rm = TRUE)),
  xlab = "Week", ylab = "Unit sales (in thousands)")
points(Data2$week, Data2$SalesPred, type = "b",
  pch = 23, col = "red", lty = 2)
```

```{r}
# generate prediction of diffusion
Data2 = Data2 %>% mutate(CumSalesPred = cumsum(SalesPred))
plot(Data2$week, Data2$CumSales, type = "p",
  ylim = c(0, 1.2 * max(Data2$CumSales, Data2$CumSalesPred, na.rm = TRUE)),
  xlab = "Week", ylab = "Cumulative sales (in thousands)")
points(Data2$week, Data2$CumSalesPred, type = "b",
pch = 23, col = "red", lty = 2)
```

p = 0.1328, q = 0.1690, and m = 1710 (in thousands). See the graphs above in the code for the sales prediction based on the p, q, and m estimates.

From the estimates and the cumulative sales prediction, the current sales have approached the saturation level. This means that no large additional increase in sales is expected without feeding gamers with extra benefits. So, it might be a good time to come up with a next movement, such as introducing an expansion pack, a new DLC (downloadable content), etc., to lift the market potential by enhancing benefits of the game.








