---
title: "Marketing W1 In Class"
author: "Carter Rudek"
date: "`r Sys.Date()`"
output: html_document
---



### Activity 2 lab material – Linear regression: GoodBelly

* This exercise is based on H.-S. Ahn (2012) “GoodBelly: Using Statistics to Justify the Marketing Expense.”
* If you’ve not yet installed R, RStudio, and tidyverse package, install them by following directions in Application installation guide for BUSML7247.pdf file.
* In this material, what you have to type in R is highlighted in yellow.

```{r}
library(tidyverse)
```

2. Data preparation

2.1. Read data (GoodBellyData.csv, available in Carmen) by using read_csv function as
follows.

```{r}
Data = read_csv("GoodBellyData.csv")
```

* If you cannot read the data, check if the working directory is the same as the folder where you’ve saved the data file by using getwd function and correct the working directory by using setwd function.

2.2. Take a look at the data frame by typing the name of the data frame (Data) as follows. This dataset has 12 variables as below. The details of the variables are available in the case article.

```{r}
Data
```


3. Basic model

3.1. Regress Sales on Price, Rep, Endcap, Demo, Demo1_3, and Demo4_5:

```{r}
reg_out = lm(Sales ~ Price + Rep + Endcap + Demo + Demo1_3 + Demo4_5, Data)
summary(reg_out)
```

3.2. Draw the residual plot of the regression for all observations:

```{r}
reg_res = resid(reg_out)
plot(reg_res, main = "Residual plot: all observations",
  xlab = "observation", ylab = "residual", col = "red")
abline(h = 0, lty = 2, lwd = 2)
```

3.3. Draw the plot of residuals grouped by Region:

```{r}
 Res = tibble(no = c(1:length(reg_res)),
  Region = Data$Region,
  residual = reg_res)
 ggplot(data = Res) +
  geom_point(mapping = aes(x = no, y = residual, color = Region))
```

* To draw a residual plot, we first need to generate a vector of residual values. resid function generates the residual vector from a regression result variable (reg_out in this exercise). See R help (type ? resid to see the help) for the details of the function.

* To draw a plot of data with grouping, ggplot (a component of tidyverse package) is much more useful than the plain function in R. See Chapter 1 of “R for Data Science” book (Chapter 3 in the web version) for the details of the function.

* Before you draw the plot by using ggplot, you need to have a data.table of what you want to plot: x-axis of the plot is the number of each observation (you can generate this simply by c(1:length(reg_res))), y-axis of the plot is the residual values (they are stored in reg_res), and the grouping variable (you can use Region in Data, i.e., Data$Region). tibble function is used to generate a new data.table. Refer to Chapter 7 of “R for Data Science” book (Chapter 10 in the web version) for the details of the function.

* What we want to draw is a scatter plot. geom_point function is combined with the ggplot function to draw the scatter plot. ggplot uses + operator instead of %>% operator to pipe functions. See Chapter 1 of “R for Data Science” book (Chapter 3 in the web version) for the details of the function.


























