---
title: "Marketing W1 In Class"
author: "Carter Rudek"
date: "`r Sys.Date()`"
output: html_document
---



### Activity 2 lab material – Linear regression: GoodBelly

* This exercise is based on H.-S. Ahn (2012) “GoodBelly: Using Statistics to Justify the Marketing Expense.”
* If you’ve not yet installed R, RStudio, and tidyverse package, install them by following directions in Application installation guide for BUSML7247.pdf file.
* In this material, what you have to type in R is highlighted in yellow.

```{r}
library(tidyverse)
```

2. Data preparation

2.1. Read data (GoodBellyData.csv, available in Carmen) by using read_csv function as
follows.

```{r}
Data = read_csv("GoodBellyData.csv")
```

* If you cannot read the data, check if the working directory is the same as the folder where you’ve saved the data file by using getwd function and correct the working directory by using setwd function.

2.2. Take a look at the data frame by typing the name of the data frame (Data) as follows. This dataset has 12 variables as below. The details of the variables are available in the case article.

```{r}
Data
```


3. Basic model

3.1. Regress Sales on Price, Rep, Endcap, Demo, Demo1_3, and Demo4_5:

```{r}
reg_out = lm(Sales ~ Price + Rep + Endcap + Demo + Demo1_3 + Demo4_5, Data)
summary(reg_out)
```

3.2. Draw the residual plot of the regression for all observations:

```{r}
reg_res = resid(reg_out)
plot(reg_res, main = "Residual plot: all observations",
  xlab = "observation", ylab = "residual", col = "red")
abline(h = 0, lty = 2, lwd = 2)
```

3.3. Draw the plot of residuals grouped by Region:

```{r}
 Res = tibble(no = c(1:length(reg_res)),
  Region = Data$Region,
  residual = reg_res)
 ggplot(data = Res) +
  geom_point(mapping = aes(x = no, y = residual, color = Region))
```

* To draw a residual plot, we first need to generate a vector of residual values. resid function generates the residual vector from a regression result variable (reg_out in this exercise). See R help (type ? resid to see the help) for the details of the function.

* To draw a plot of data with grouping, ggplot (a component of tidyverse package) is much more useful than the plain function in R. See Chapter 1 of “R for Data Science” book (Chapter 3 in the web version) for the details of the function.

* Before you draw the plot by using ggplot, you need to have a data.table of what you want to plot: x-axis of the plot is the number of each observation (you can generate this simply by c(1:length(reg_res))), y-axis of the plot is the residual values (they are stored in reg_res), and the grouping variable (you can use Region in Data, i.e., Data$Region). tibble function is used to generate a new data.table. Refer to Chapter 7 of “R for Data Science” book (Chapter 10 in the web version) for the details of the function.

* What we want to draw is a scatter plot. geom_point function is combined with the ggplot function to draw the scatter plot. ggplot uses + operator instead of %>% operator to pipe functions. See Chapter 1 of “R for Data Science” book (Chapter 3 in the web version) for the details of the function.


### Question and Answers


###### Q1. Interpret the following coefficient estimates of the regression model in the lab. (For example, what does price coefficient = -28.609 mean?)

* Price = -28.609

With a $1 increase in the average retail price, the weekly demand decreased by about 29 units.

* Rep = 76.951

Sales at stores with a designated sales representative were higher than at stores with no designated sales rep (the different is, on average, about 77 units).

* Endcap = 304.960

Weekly average sales increased by about 305 units per week during the weeks of the endcap competition.

* Demo = 111.261

The average sales increased by about 111 units in the week of the demo.

* Demo1_3 = 73.663 

The average sales increased by about 74 units if the demo took place between one and three weeks ago.

* Demo4_5 = 67.700

The average sales increased by about 68 units if the demo took place four weeks ago or more. 

* Based on the interpretation, what's your conclusions on the effects of three promotion activities (in-store demo, face-to-face sales representative, and end-cap display)?

The coefficient estimates and their p-values support that all of those activities boost the sales.



###### Q2. In the residual plot of the lab, you may find that some outliers exist in specific regions. This suggests an inclusion of the region effects. In the dataset, we have the region variable (Region), which is a categorical variable. Add this variable into the regression model as an additional regressor and run the regression again, as follows:

* When a categorical variable is added into a regression, it should be designated as a factor-type variable by using as.factor function, i.e., lm(Sales ~ Price + Rep + Endcap + Demo + Demo1_3 + Demo4_5 + as.factor(Region), Data).

Does the inclusion of the region effects make changes in the effects you've found in Q1?

```{r}
#Code and results:
reg_out = lm(Sales ~ Price + Rep + Endcap + Demo + Demo1_3 + Demo4_5 +
as.factor(Region), Data)
summary(reg_out)
```

```{r}
reg_res = resid(reg_out)
plot(reg_res, xlab = "observation", ylab = "residual")
abline(h = 0, col = "gray")
```

Variables             Model w/o Region            Model w/ Region
Price                     -28.609                    -33.114
Rep                        76.951                     35.127
Endcap                     304.960                    302.020
Demo                       111.261                    113.056
Demo1_3                    73.663                     74.044
Demo4_5                    67.700                     65.957


The table above summarizes coefficient estimates of the price and promotion variables for both the model w/o Region variables and the model w/ Region variables. With consideration of standard errors, the inclusion of Region variables affects only the sales representative promotion effect (76.951 -> 35.127). It does not affect all the other effects (the difference in coefficient estimates is relatively small, e.g., 111.261 -> 113.056 for Demo variable).


















