---
title: "TS Forecasting Final Project - Group 12"
author: "Carter Rudek, Brandon Slover & Chris Garcia"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

# Packages
library(pacman)
pacman::p_load(keyring, openxlsx, DBI, odbc, readr, dbplyr, slider, GGally, seasonal, fpp3, gridExtra, htmltools, devtools, feasts, tidyverse, lubridate)
```

```{r Read Data}
df <- read.csv("forecasting_data.csv")
cpi <- read.csv("cpi_2015_Mar2023.csv")
```


-------------------------------------------------------------------------------------------------------------------------------

# 1. Data Preparation

### Sales (Dependent Variable)

```{r Sales Preparation}
# add month column
df <- df %>%
  mutate(date = mdy(TRX_DATE))

df <- df %>%
  mutate(month = yearmonth(lubridate::floor_date(df$date, "month")))

# create ts, filter data from 2016 through March 2023
df_ts <- df %>%
  filter(between(date, as.Date("2016-01-01"), as.Date("2023-03-31"))) %>% #change this once we get April data
  select(month, NET_SALES) %>%
  group_by(month) %>%
  summarise(net_sales = sum(NET_SALES)) %>%
  as_tsibble(index = month)
```


### Regression data (Independent Variables)

##### Consumer Price Index (CPI)

```{r CPI}
# clean CPI and create ts
cpi_ts <- cpi %>%
  filter(year>=2016) %>%
  select(year, periodName, value) %>%
  unite(my, c("periodName", toString("year"))) %>%
  mutate(my = str_replace(my,"_"," ")) %>%
  mutate(date = my(my)) %>%
  mutate(month = yearmonth(date, "month")) %>%
  select(month, value) %>%
  rename(cpi = 'value') %>%
  as_tsibble(index = month)

# join sales and cpi data
sales_cpi_ts <- df_ts %>%
  full_join(cpi_ts, by = "month")
```

##### Covid

```{r}
# add covid dummy variable to data for regression
sales_cpi_covid_ts <- sales_cpi_ts %>%
  mutate(covid = ifelse(
    time_in(month, "2020 Jan" ~ "Dec 2021"),
    1, 0
  ))
```


-------------------------------------------------------------------------------------------------------------------------------

# 2. Firm Information







-------------------------------------------------------------------------------------------------------------------------------

# 3.Technique Application



### i. Descriptive Visualization

```{r}
# 1. continuous line chart 
options(scipen = 999)
df_ts %>%
  ggplot(aes(label = month, x = month, y = net_sales, color = "red")) +
  geom_line() +
  geom_smooth(method = "lm") +
  geom_point() +
  geom_vline(xintercept = as.Date("2020-03-01"), color = "grey", linetype="dotted") + # covid start?
  geom_vline(xintercept = as.Date("2021-12-01"), color = "grey", linetype="dotted") + # covid end?
  geom_point(data = df_ts[which.min(df_ts$net_sales), ], color="blue", 
             size=3) +
  geom_point(data = df_ts[which.max(df_ts$net_sales), ], color="green", 
             size=3) +
  geom_text(data = bind_rows(df_ts[which.min(df$net_sales), ], 
                             df_ts[which.max(df$net_sales), ]), 
            aes(label = month)) +
  scale_y_continuous(labels = scales::comma) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = 'black'),
    plot.background = element_rect(fill = 'black'),
    text=element_text(color="white"),axis.text=element_text(color="white")) + 
  labs(title='Monthly Sales') +
  xlab("Time") +
  ylab("Net Sales") +
  theme(legend.position="none")
```

```{r}
# 2. seasonal plot by year
options(scipen = 999)
df_ts %>% 
  gg_season(net_sales, labels = "both") +
  scale_y_continuous(labels = scales::comma) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = 'black'),
    plot.background = element_rect(fill = 'black'),
    text=element_text(color="white"),axis.text=element_text(color="white"))+ 
  labs(title='YoY by Month') +
  xlab("Time") +
  ylab("Net Sales")
```

```{r}
# 3. seasonal plot by month
options(scipen = 999)
df_ts %>% 
  gg_subseries(net_sales, color = "red") +
  scale_y_continuous(labels = scales::comma) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_rect(fill = 'black')) +
  ggtitle("Seasonal Plot by Month") +
  xlab("Years") +
  ylab("Net Sales")
```

```{r}
# 4. ACF
df_ts %>%
  ACF(net_sales, lag_max = 48) %>%
  autoplot() +
  labs(title = "Net Sales")
```

```{r}
# 5. Decomposition?

Placeholder
Placeholder
Placeholder

```


```{r CPI and Sales}
# 6. Visualization/Exploration for dependent and independent variable
## all
sales_cpi_ts %>%
  pivot_longer(!month, names_to = "metric", values_to = "number") %>%
  ggplot(aes(x = month, y = number, group = metric)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_grid(vars(metric), scales = "free_y") +
  ggtitle('Full Date Range')

## investigate prior to covid
sales_cpi_ts %>%
  filter_index(.~ "2019 Dec") %>%
  pivot_longer(!month, names_to = "metric", values_to = "number") %>%
  ggplot(aes(x = month, y = number, group = metric)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_grid(vars(metric), scales = "free_y") +
  ggtitle('Prior to Covid')

## investigate covid to present
sales_cpi_ts %>%
  filter_index("2020 Jan" ~.) %>%
  pivot_longer(!month, names_to = "metric", values_to = "number") %>%
  ggplot(aes(x = month, y = number, group = metric)) +
  geom_line() +
  geom_smooth(method = "lm") +
  facet_grid(vars(metric), scales = "free_y") +
  ggtitle('Covid-Present')

```

### ii. ARIMA / Dynamic Regression



# Statistical Modeling


```{r}
# basically no correlation
cor(sales_cpi_covid_ts$net_sales, sales_cpi_covid_ts$cpi)

# scatter matrix
sales_cpi_covid_ts %>% 
  ggpairs(col=c(2:3))

```


### Regression


```{r linear model}
# Model with all potential variables
# fit TSLM
lm <- sales_cpi_covid_ts %>% 
  model(tslm = TSLM(net_sales ~ cpi + as.factor(covid) + lag(net_sales,1) + lag(net_sales,2) + lag(net_sales,12) + trend() + season()))

#report lm output
report(lm)
print("---------------------------------------------------------------------")

# fit metrics
lm  %>% glance %>%
select(adj_r_squared, AIC, AICc, BIC, CV) 
```

```{r}
# check residuals
lm %>% gg_tsresiduals()

# ljung box test for autocorrealtion
augment(lm) %>%
  features(.resid, ljung_box, lag = 10, dof = 8)
```

```{r}
# Compare multiple models

#fit several models with different lags and see which is best (note use of differenced variables)
fit1 <- sales_cpi_covid_ts %>% model(
model1 = TSLM(net_sales ~ cpi),
model2 = TSLM(net_sales ~ as.factor(covid)),
model3 = TSLM(net_sales ~ cpi + as.factor(covid)),
model4 = TSLM(net_sales ~ cpi + as.factor(covid) + lag(net_sales,1)),
model5 = TSLM(net_sales ~ cpi + as.factor(covid) + lag(net_sales,1) + lag(net_sales,2)),
model6 = TSLM(net_sales ~ cpi + as.factor(covid) + lag(net_sales,1) + lag(net_sales,2) + lag(net_sales,12)),
model7 = TSLM(net_sales ~ cpi + as.factor(covid) + lag(net_sales,1) + lag(net_sales,2) + lag(net_sales,12) + trend()),
model8 = TSLM(net_sales ~ cpi + as.factor(covid) + lag(net_sales,1) + lag(net_sales,2) + lag(net_sales,12) + trend() + season()),
model9 = TSLM(net_sales ~ cpi + as.factor(covid) + trend() + season())
)
fit1 %>% glance %>% select(adj_r_squared, AIC, AICc, BIC, CV) 

```

```{r}
# example forecasting with just cpi
cpi_model <- sales_cpi_covid_ts %>% 
  model(tslm = TSLM(net_sales ~ cpi))

#set number of periods to forecast
h=12

#create newdata for first scenario
scen_data <- scenarios(
  "avg_cpi" = new_data(sales_cpi_covid_ts, h) %>% 
    mutate(
      cpi = rep(mean(sales_cpi_covid_ts$cpi),h), # average p
    names_to = "Scenario"))


cpi_mod_avg <- forecast(cpi_model, new_data = scen_data)

#create plot
sales_cpi_covid_ts %>%
  autoplot(net_sales) +
  autolayer(cpi_mod_avg) +
  ylab('Log Units Sold') +
  xlab('Time (in Weeks)') +
  ggtitle('Forecast for Log Units Sold of Tropicana Pure Premium 64oz using Avergage Price and Coupon/D')
```

### ARIMA

```{r}
# check for stationarity
sales_cpi_covid_ts %>% 
  features(net_sales, c(unitroot_kpss, unitroot_ndiffs, unitroot_nsdiffs))
```

It doesn't seem like anything other than a seasonal diff needs done, let's go through some of the transformations

```{r}
#original plot
plot1 <- sales_cpi_covid_ts %>% 
  select(month, net_sales) %>%
  autoplot()+ #create plot of time series
  xlab("Time")+ylab("Value")+ #change labels
  ggtitle("Time Series Plot 1")

plot1
```

BOX-COX for unstable variance (fluctuations)

```{r}
# There doesnt seem to be any problem with unstable variance

# look at what box cox does to the data
optlambda <- sales_cpi_covid_ts %>% 
  select(month, net_sales) %>%
  features(net_sales, features=guerrero) %>% #get optimal lambda and pull value
  pull(lambda_guerrero)

bc_trans <- sales_cpi_covid_ts %>% 
  select(month, net_sales) %>%
  mutate(net_sales = box_cox(net_sales, optlambda)) #create transformed time series

plot2 <- bc_trans %>% autoplot() + #create plot of transformed series
  xlab("Time")+ylab("BoxCox Times Series")+ #change axis labels
  ggtitle("Box Cox Transformed Values")

grid.arrange(plot1, plot2, ncol=1) #combine plots of time series
```

First Difference to handle trends


```{r}
# there isnt much trend, this doesnt do much either

# look what first differencing does
fd_trans <- sales_cpi_covid_ts %>% 
  select(month, net_sales) %>% 
  mutate(net_sales = difference(net_sales,1)) #create differenced series

plot3 <- fd_trans %>% autoplot() + #create plot of diff’d transformed series
  xlab("Time")+ylab("Diff'd Box Cox Values")+ #change axis labels
  ggtitle("First Differenced Box Cox Transformed Values")
grid.arrange(plot1, plot3, ncol=1) #combine plots of time series
```

Seasonal Difference

```{r}
sd_trans <- sales_cpi_covid_ts %>% 
  select(month, net_sales) %>%  
  mutate(net_sales = difference(net_sales,12)) #create seasonally differenced series


plot4 <- sd_trans %>% autoplot() + #create plot of diff’d series
  xlab("Time")+ylab("Seasonally Diff'd Values")+ #change axis labels
  ggtitle("Seasonally Differenced Series")
grid.arrange(plot1, plot4, ncol=1) #combine plots of time series
```


```{r}
# check for stationarity for series with seasonal difference
sd_trans %>% 
  features(net_sales, c(unitroot_kpss, unitroot_ndiffs, unitroot_nsdiffs))
```

Everything looks good now!

```{r}
#plot ACF and PACF plots with series
sd_trans %>% 
  gg_tsdisplay(net_sales, 'partial')
```

An AR(1) model:

* Has an ACF plot that decays gradually over many lags
* Has a PACF plot that decays immediately after a single lag

We took a seasonal difference 

Seasonal MA(1)
* exponential decay in the seasonal lags of the ACF;
* a single significant spike at lag 12 in the PACF.

```{r}
# the arima model i am trying based on acf and pacf plots
my_arima_fit <- sales_cpi_covid_ts %>% 
  select(month, net_sales) %>%
  model(
    ARIMA(net_sales ~ pdq(1,0,0) + PDQ(0,1,1))
    )

report(my_arima_fit)
```










##### Automatically have R generate

```{r}
# ORIGINAL
# ask R to identify the best fitting ARIMA model
arima_fit <- sales_cpi_covid_ts %>% 
  select(month, net_sales) %>%
  model(
    ARIMA(net_sales)
    )

report(arima_fit)
```

```{r}
# New Seasonally Differenced data
# ask R to identify the best fitting ARIMA model
arima_sd_fit <- sd_trans %>% model(
  ARIMA(net_sales))

report(arima_sd_fit)
```

```{r}
sales_cpi_covid_ts %>% model(ARIMA(net_sales ~ pdq(1,0,0) + PDQ(0,1,1))) %>%
  forecast(h=12) %>%
  autoplot(sales_cpi_covid_ts)
```
























### Dynamic Regression









### iii. Forecast






















