---
title: "BUSOBA 6400: Assignment Week 2 Data Viz/Summary"
author: "Carter Rudek"
output: html_document
---


```{r, echo = FALSE, message = FALSE, warning = FALSE, results = FALSE}
#load packages
library(tidytuesdayR)
library(tidyverse)
library(rmarkdown)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, results = FALSE}
#read data
tuesdata <- tidytuesdayR::tt_load('2022-07-05')
tuesdata <- tidytuesdayR::tt_load(2022, week = 27)

rent <- tuesdata$rent
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.width=10}
rent %>%
  filter(year == 2018, county != "NA") %>%
  ggplot(aes(x=county, y=price)) +
  geom_boxplot() +
  geom_jitter(width = .35, alpha=.12) +
  ggtitle("San Francisco Rent Prices by County in 2018")
```

There are a lot of properties for rent in San Francisco across the 10 counties. To make the boxplots easier to read, i jittered the points and adjusted the opacity. From these boxplots, we can learn alot. We can easily see counties with lots of rentals (Alameda, Contra Costa, San Francisco, Santa Clara) and counties with little rentals (Napa). We can see that all counties, except maybe Solano, have outliers. We can see that the median house prices in each county are relatively close, however, Marin and San Frnaciscos IQR's are larger than the other counties.