---
title: "BUSOBA 7332 Homework 7"
author: "Carter Rudek"
date: "`r Sys.Date()`"
output: html_document
---

#### Load Packages

```{r, error = FALSE}
library(arules)
library(arulesViz)
library(tidyverse)
```

### Read data

```{r}
data<-read.transactions("Bread_Basket.csv",format = "single",sep=",",header=TRUE,cols = 1:2)
```


```{r}
View(data)
```

```{r}
inspect(head(data,3))
```

#### 1. How many unique transactions (carts) were made based on the data? 

```{r}
length(data)
```

#### 2. How many unique items were there on the data?  

```{r}
length(itemFrequency(data))
```

#### 3. What are the top 2 items based on the transaction data? 

```{r}
itemFrequencyPlot(data,topN=10,type="absolute", main="Top 10 Sold Items")
```

#### 4. What percentage of all carts contain coffee?  

```{r}
sort(itemFrequency(data),decreasing = TRUE)[1:10] 
```

#### 5. Generate association rules with a minimum support of 0.0012 and minimum confidence of 0.80 and minlen=2. Sort the rules by lift ratio. Which of the following is the top rule? 

```{r}
mrules2<-apriori(data,parameter=list(supp=0.0012, confi=0.8, minlen=2)) # Finding association rules
mrules2 # How many rules
```

```{r}
inspect(sort(mrules2,by="lift",decreasing=TRUE)[1:6])  # Sort by lift 
```



#### 6. What is the lift ratio for the top rule in the previous question? Report your answer to 2 decimal places.








#### 7. Generate association rules with a minimum support of 0.0012 and minimum confidence of 0.80 and minlen=2. Sort the rules by support. What is the support for the top rule? Report your answer to 4 decimal places.

```{r}
inspect(sort(mrules2,by="support",decreasing=TRUE)[1:6]) 
```













